\section{The Structure Of Quantity Expressions}
\label{sec:strucqe}

The first step in developing a good search engine for Quantity Expressions is to take a closer look at Quantity Expressions.

When looking at Quantity Expressions, we need to distinuish between two perspectives, a notational one (i. e. the way they are written down) and a semantic one (i. e. what they mean).

\subsection{Compositional Behaviour Of Quantity Expressions}
Let us start by the behaviour of the notation. For this purpose let us take a look at:
\[x = 25\ \frac{\text{m}}{\text{s}}\]
We notice that $x$ consists of 2 parts, a scalar ($25$) and a scalar-free $\frac{\text{m}}{\text{s}}$ one. Furthermore, the unit consists of two primitive units $m$ and $s$. Since they divide one another, we can conclude that $x$ describes a velocity.

In general, the \textit{unit component} is not always scalar free. While $m$ and $s$ certainly do not contain scalars, this is no longer the case in the following example:
\[y = \frac{\text{L}}{100\ \text{km}}\]
We could certainly call $y$ a \textit{unit}. It consists of 2 sub-expressions, $\text{L}$ and $100\ \text{km}$. The first one is a primitive unit and the second one a multiplication of a number and the unit $\text{km}$. This, first of all, leads us to define the concept of a \textit{primitive unit} as a simple Quantity that is not a Number.

Abstracting from the examples above, we can now considder the semantics of Quantity Expressions. It makes sense to define Quantity Expressions as one of the following 6 cases: 
\begin{enumerate}
  \item A \textit{primitive unit}, such as \textit{Meter}.
  \item The \textit{Product} of a Quantity Expression with a scalar, such as $5 \text{Meter}$
  \item The \textit{Quotient} of a Quantity Expression by a (non-zero) scalar\footnote{Strictly speaking, this is equivalent to the case above. For simplicity we define it anyways. }.
  \item The Multiplication $\cdot{}$ which takes 2 existing quantity expression and generates a new one, for example $\cdot \left(100, \text{m} \right) = 100\ \text{m}$
  \item The division / which again takes 2 quantity expressions and generates a new one, for example $\text{/} (\text{m}, \text{s}) = \frac{\text{m}}{\text{s}}$
  \item The addition of two quantity expressions
\end{enumerate}

This allows us to easily generate Quantity Expressions (like the examples $x$ and $y$ from above) starting at the primitive units m, s, L and km.

\subsection{Dimensions Of Quantity Expressions}
Next let us briefly examine the dimensions of Quantity Expressions. The dimension of a Quantity Expression is the type of quantity it expresses. For example $5\ \text{m}$ describes some length. According to the International System of Units \cite{sispec} there are seven basic dimensions:
\begin{enumerate}
  \item length
  \item mass
  \item time
  \item electric current
  \item temperature
  \item luminous intensity
  \item amount of substance.
\end{enumerate}

In addition to these we found that there are commonly two other dimensions: The \textit{count} dimension (used for counting of objects) and the \textit{none} dimension for dimensionless quantities.

Similar to the compositional behaviour of Quantity Expressions, dimensions can be multiplied and divided. In this the \textit{count} dimension behaves like a multiplicative identity element.

Unlike Quantity Expressions however dimensions can not be multiplied with numbers. Furthermore, when multiplying to Quantity Expressions of dimensions $a$ and $b$ their resulting dimension is $a \cdot{} b$, the product of the dimensions. The same goes for division. In this regard the dimension of a quantity expression behaves like a type.

\subsection{A Mathematical Theory Of Quantity Expressions}
\label{sec:qeform}

The realisation that dimensions are types of quantity expressions leads us to our first formalisation of Quantity Expressions. We first define a theory of dimensions in Figure \ref{fig:dimensions} and then import it to define a theory of Quantity Expressions in Figure \ref{fig:QE}.

\input{graphs/dimensions.tex}

\input{graphs/QE.tex}

Figure \ref{fig:dimensions} defines the 9 basic dimensions and then dimension compositon via multiplication and division. Then we move on in Figure \ref{fig:QE} to define quantity expressions. Each quantity expression has a dimension (via the \textit{QE} constant). This allows us to define basic units (which we will actually do in the next Figure). With the \textit{QENMul} and \textit{QENDiv} symbols we can multiply and divide quantity expressions by numbers (for this case we actually need to import some theory of numbers to allow us actually write this down as a Term). Then we define multiplication and division of quantity expressions in such a way that dimensions multiply and divide appropriatly.

Now we need to introduce some basic units. Let us start by just defining \textit{Meter} in Figure \ref{fig:meter}. We can now write a term in this theory that expresses any number of meters.

\input{graphs/meter.tex}

\subsection{Transforming Quantity Expressions From One Form Into Another}
\label{sec:qeconv}

This provides a basis for a unit system. Let us furthermore define a few more units of non-metric lengths. In Figure \ref{fig:nonsilength} we show a few non-si units. Here we first define thoug as a quanity expression of length and then one-by-one define more units in terms of the previous one.
\input{graphs/nonsilength.tex}

We now want to relate quantity expressions with units from the \textit{Meter} theory to units from the \textit{Non SI Lengths} theory. It is known that $1 \mathsf{Thou} = 0.0000254 \mathsf{Meter}$. This can be easily expressed with a view $\psi$ between these two theories:

\[
\psi = \left\{\begin{array}{l}
  \mathsf{Thou} \mapsto \mathsf{QENMul} \left( 0.0000254, \mathsf{Meter} \right)
\end{array}\right\}
\]

Even though the view just maps the symbol \textit{Thou} to some Term in the Meter theory, we can also use it to transform any other term from the Thou Theory. Since all units are defined in terms of the previous one, we can just expand all definitions to get an expression containing only numbers and the unit \textit{Thou}. Then we can use the view as normal to get a Quantity Expression in the \textit{Meter} theory.
